Make any stupid plot you want

### 2021-04-16 draft/sections/math.tex
graphics section

As illustrated in figure~\ref{fig:math:data:fiber}, Spivak's framework provides a consistent way to describe potentially complex components of the input data.

@@

The datasets in figure~\ref{fig:math:data:base} have the same fiber of (temperature, time). The dot represents a discrete base space \dbase, meaning that every dataset encoded in the fiber bundle has discrete continuity. The line is a representation of a 1D continuity, meaning that every dataset in the fiber bundle is 1D continous. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. The explicit topology is a concise way of distinguishing visualizations that appear identical, for example heatmaps and images.

@@

@@

\subsection{Graphic Space \gtotal}
\label{sec:graphic}  
We introduce a graphic bundle to hold the essential information necessary to render a graphical design constructed by the artist. As with the data, we can represent the target graphic as a section \gsection\ of a bundle  $(\gtotal, \gbase, \pi, \gfiber)$. The graphic bundle \gtotal\ consists of a base \gbase (~\ref{sec:graphic_fiber}) that is a thickened form of \dbase\, 
a fiber \gfiber (~\ref{sec:graphic_base}) that is an idealized display space, and sections \gsection (~\ref{sec:graphic_section}) that encode a graphic where the visual characteristics are fully specified.

\subsubsection{Idealized Display \gfiber}
\label{sec:graphic_fiber}
To fully specify the visual characteristics of the image, we construct a fiber \gfiber\ that is an infinite resolution version of the target space. Typically \gtotal\ is trivial and therefore sections can be thought of as mappings into \gfiber. In this work, we assume a 2D opaque image $\gfiber=\reals^5$ with elements 
\begin{equation}
(x,\, y,\, r,\, g,\, b) \in \gfiber
\end{equation}
such that a rendered graphic only consists of 2D position and color. To support overplotting and transparency, the fiber could be $\gfiber=\reals^{7}$ such that $(x, y, z, r, g, b, a) \in \gfiber$ specifies the target display. By abstracting the target display space as \gfiber, the model can support different targets, such as a 2D screen or 3D printer.

\subsubsection{Continuity of the Graphic \gbase} 
\label{sec:graphic_base}
Just as the \dbase\ encodes the connectivity of the records in the data, we propose an equivalent \gbase\ that encodes the connectivity of the rendered elements of the graphic. For example, consider a \gbase\ that is mapped to the region of a 2D display space that represents \dbase. For some visualizations, \dbase\ may be lower dimension than \gbase. For example, a point that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D to encode the connectivity of the pixels that visually represent the point. This thickening is often not necessary when the dimensionality of \dbase\ matches the dimensionality of the target space, for example if \dbase\ is 2D and the display is a 2D screen. We introduce \gbase\ to thicken \dbase\ in a way which preserves the structure of \dbase.

Formally, we require that \dbase\ be a deformation retract\cite{RetractionTopology2020} of \gbase\ so that \dbase\ and \gbase\ have the same homotopy. The surjective map $\vindex: \gbase \rightarrow \dbase$ 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
    \end{tikzcd}
\end{equation}
goes from region $\gbasepoint \in \gbase_{\dbasepoint}$ to its associated point $\gbasepoint$. This means that if $\vindex(\gbasepoint) = \dbasepoint$, the record at \dbasepoint\ is copied over the region \gbasepoint\ such that $\dsection(\dbasepoint)=\vindex^*\dsection(\gbasepoint)$ where $\vindex^*\dsection(\gbasepoint)$  is \dsection\ pulled back over \gbase.

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/retraction_maps.png}
    \caption{The scatter and line graphic base spaces have one more dimension of continuity than \dbase\ so that \gbase\ can encode physical aspects of the glyph, such as shape (a circle) or thickness. The image has the same dimension in \gbase\ as in \dbase.}
    \label{fig:graphic_retraction_map}
\end{figure}

When \dbase\ is discrete points and the graphic is a scatter plot, each point $\dbasepoint \in \dbase$ corresponds to a 2D disk $\gbase_{\dbasepoint}$ as shown in figure~\ref{fig:graphic_retraction_map}. In the case of 1D continuous data and a line plot, the region \gy\ over a point $\gx_i$ specifies the thickness of the line in \gbase\ for the corresponding \dsection\ on \dbasepoint. The image has the same dimensions in data space and graphic space such that no extra dimensions are needed in \gbase.

@@

\subsubsection{Graphic \gsection}
\label{sec:graphic_section}
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/render.png}
    \caption{To render a graphic, a pixel $p$ is selected in the display space, which is defined in the same coordinates as the x and y components in \gfiber.  The inverse mapping ${\gsection_{xy}}^(p)$ returns a region $\gbase_{p} \subset \gbase$. $\gsection(\gbase_{p})$ returns the list of elements $(x,\,y,\,r,\,g,\,b) \in \gfiber$ that lie over $\gbase_{p}$. The integral over the $(r,\,g,\,b)$ elements is the color of the pixel.}
    \label{fig:graphic_rho_lookup}
\end{figure}

This section describes how we go from a graphic in an idealized prerender space to a rendered image, where the graphic is the section $\gsection: \gbase \rightarrow \gtotal$. It is sufficient to sketch out how an arbitrary pixel would be rendered, where a pixel $p$ in a real display corresponds to a region $\gbase_p$ in the idealized display. To determine the color of the pixel, we aggregate the color values over the region via integration.

For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{xy}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. To compute the color, we integrate on $\gbase_p$

\begin{align}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align}

As shown in figure~\ref{fig:graphic_rho_lookup}, a pixel $p$ in the output space is selected and inverse mapped into the corresponding region $\gbase_p \subset \gbase$. This triggers a lookup of the $\gsection$ over the region $\gbase_p$, which yields the set of elements in \gfiber\ that specify the $(r, g, b)$ values corresponding to the region $p$. The color of the pixel is then obtained by taking the integral of $\gsection_{rgb}(\gbase_p)$.

In general, \gsection\ is an abstraction of rendering. In very broad strokes \gsection\  can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, or an openGL scene graph\cite{CarsonOpenGL1997}. Alternatively, \gsection can be a rendering engine such as cairo\cite{CairographicsOrg} or AGG\cite{AntiGrainGeometry}. Implementation of \gsection\ is out of scope for this work,

\subsection{Artist}
\label{sec:artist}
We propose that the transformation from data to visual representation can be described as a structure preserving map from one topological space to another. We name this map the artist as that is the analogous part of the  Matplotlib\cite{hunterArchitectureOpenSource} architecture that builds visual elements. The topological artist \vartist\ is a monoid equivariant sheaf map from the sheaf on a data bundle \dtotal\ which is $\mathcal{O}(\dtotal)$ to the sheaf on the graphic bundle \gtotal, $\mathcal{O}(\gtotal)$. 
\begin{equation}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation}
Sheafs are a mathematical object with restriction maps that define how to glue \dsection over local neighborhoods $U \subseteq \dbase$,  discussed in section~\ref{sec:artist_sec:data_sheaf_stalk}, such that the \vartist\ maps are consistent over continuous regions of \dbase. While \vartist can usually construct graphical elements solely with the data in \dsection, some visualizations, such as line, may also need some finite number $n$ of derivatives, which is captured by the jet bundle $\mathcal{J}^n$ \cite{JetBundle2020,musilovaCalculusVariationsJet2016} with $\mathcal{J}^{0}(\dtotal)=\dtotal$. In this work, we at most need $\mathcal{J}^{2}(\dtotal)$ which is the value at \dsection\ and its first and second derivatives; therefore the artist takes as input the jet bundle $\dtotal^{\prime}=\mathcal{J}^{2}(\dtotal)$.

@@

\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
                                              & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
        \end{tikzcd}
\end{equation}
where the input can be point wise $\dsection(\dbasepoint)\mid \dbasepoint \in \dbase$. The encoders $\vchannel:\dtotal^{\prime} \rightarrow \vtotal$ convert the data components to visual components(\ref{sec:graphic_base}). The continuity map $\vindex:\gbase \rightarrow \dbase$ then pulls back the visual bundle \vtotal\ over \gbase (\ref{sec:artist_nu}). Then the assembly function $\vmark: \vtotalpull \rightarrow \gtotal$ composites the fiber components of \vtotalpull\ into a graphic in \gtotal (\ref{sec:artist_q}). This functional decomposition of the visualization artist facilitates building reusable components at each stage of the transformation because the equivariance constraints are defined on \vchannel, \vmark, and \vindex.

\subsubsection {Visual Fiber Bundle \vtotal}
We introduce a visual bundle \vtotal\ to store the visual representations the artist needs to assemble into a graphic. The visual bundle (\vtotal, \dbase, $\pi$, \vfiber) has section $\vsection: \vtotal \rightarrow \dbase$ that resolves to a visual variable in the fiber \vfiber. The visual bundle \vtotal\ is the latent space of possible parameters of a visualization type, such as a scatter or line plot. We define \vfiber\ in terms of the parameters of a visualization libraries compositing functions; for example table~\ref{tab:mpl_visual_variable_fiber} is a sample of the fiber space for Matplotlib \cite{hunterMatplotlib2DGraphics2007}.

@@

\subsubsection{Visual Encoders \vchannel}
\label{sec:artist_nu}
%% equivariant maps 
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/partial_fixed.png}
    \caption{In this artist, \vchannel\ maps the strings to the emojis. This \vchannel\ is equivariant because the monoid actions (which are represented by the colored arrows) are the same on both the \dsection\ input and \vsection\ output sets.}
    \label{fig:artist_nu}
\end{figure}
As introduced in section~\ref{sec:intro_viz}, there are many ways to visually represent data components. We define the visual transformers \vchannel\ 
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in section~\ref{sec:data_monoid}, we choose monoid actions as the basis for equivariance because they define the structure on the fiber components.

@@

On example of an equivariant \vchannel\ is illustrated in figure~\ref{fig:artist_nu}, which is a mapping from \textbf{Strings} to symbols. The data is an example of a Steven's nominal measurement set, which is defined as having on it permutation group actions 
\begin{equation}
    \text{ if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)
\end{equation}
such that shuffling the words must have an equivalent shuffle of the symbols they are mapped to. This is illustrated in the identical actions, represented by the colored arrows, on the words and emojis. To preserve ordinal and partial order monoid actions, \vchannel\ must be a monotonic function such that given $\delement_1, \delement_2 \in \dtotal_{i}$,
\begin{equation}
\text{ if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)
\end{equation}
the visual encodings must also have some sort of ordering. For interval scale data, \vchannel\ is equivariant under translation monoid actions if 
\begin{equation}
\vchannel (x + c) = \vchannel(x) + c
\end{equation}
while for ratio data, there must be equivalent scaling\cite{weissteinSimilarityTransformation} 
\begin{equation}
\vchannel(xc) = \vchannel(x)*c 
\end{equation}
We therefore can test if a \vchannel\ is equivariant by testing the actions under which is must  commute. For example, we define a transform $\vchannel_{i}(x) = .5$ on interval data. This means it must commute under translation, for example $t(x) = x+2$. Testing this constraint 
\begin{align}
    \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + 2\\
    .5 &\neq .5 + 2
\end{align}
we find that the \vchannel\ defined here does not commute and is therefore invalid. The constraints on \vchannel\ can be embedded into our artist such that the \vchannel\ functions can test for equivariance and also provide guidance on constructing new \vchannel\ functions.

@@

As shown in figure~\ref{fig:artist_q}, the assembly function \vmark\ combines the fiber $\dfiber_i$ wise \vchannel\ transforms into a graphic in \gtotal. Together, \vchannel\ and \vmark\ are a map-reduce operation: map the data into their visual encodings, reduce the encodings into a graphic. As with \vchannel\, the constraint on \vmark\ is that for every monoid action on the input \vsection\, there is corresponding monoid action on the output \gsection.

@@

The glyph in figure~\ref{fig:artist_mark_change} has the following characteristics \vfiber\ specified by  $(xpos,\, ypos,\, color,\, thickness)$ such that one section is $\vsection=(0,0,0,1)$ and $\vmark(\vsection) = \gsection$ generates a piece of the thin hollow circle. The equivariance constraint on \vmark\ is that the action $m=(e, e, e, x+2)$, where e is identity, translates \vsection\ to  $\vsection^{\prime}=(e,e,e,3)$. The corresponding action on \gsection\ causes $\vmark(\vsection^{\prime})$ to be the thicker circle in figure~\ref{fig:artist_mark_change}.

\subsubsection{Assembly \vmark}
In this section we formulate the minimal Q that will generate distinguishable graphical marks: non-overlapping scatter points, a non-infinitely thin line, and an image. 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/scatter_with_s.png}
    \caption{The data is discrete points (temperature, time). Via \vchannel\ these are converted to (xpos, ypos) and pulled over discrete \gbase. These values are then used to parameterize \gsection\ which returns a color based on the parameters (xpos,ypos) and position $\alpha, \beta$ on $\gbase_{\dbasepoint}$ that \gsection\ is evaluated on. 
    }
    \label{fig:artist_scatter}
\end{figure}
The scatter plot in figure~\ref{fig:artist_scatter} can be defined as $\vmark(xpos, ypos)(\alpha, \beta)$ where color $\gsection_{RGB} = (0,0,0)$ is defined as part of \vmark\ and $\gbasepoint=(\alpha, \beta)$ defines the region on \gbase. The position of this swatch of color can be computed relative to the location on the disc $\gbase_{\dbasepoint}$ as shown in figure~\ref{fig:artist_scatter}:
\begin{align}
x &= size *\alpha \cos(\beta) + xpos \\
y &= size *\alpha \sin(\beta) + ypos
\end{align}

such that $\gsection(\gbasepoint) = (x, y, 0, 0, 0)$ colors the point (x,y) black. Here $size$ can either be defined inside \vmark\ or it could also be a parameter in \vtotal\ that is passed along with (xpos, vpos). As seen in figure~\ref{fig:artist_scatter}, a scatter has a direct mapping from a region on $\gbase_{k}$ to its corresponding \dbasepoint.

>>>>

As illustrated in \autoref{fig:math:data:fiber}, Spivak's framework provides a consistent way to describe potentially complex components of the input data.

@@

The datasets in \autoref{fig:math:data:base} have the same fiber of (temperature, time). The dot represents a discrete base space \dbase, meaning that every dataset encoded in the fiber bundle has discrete continuity. The line is a representation of a 1D continuity, meaning that every dataset in the fiber bundle is 1D continous. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. The explicit topology is a concise way of distinguishing visualizations that appear identical, for example heatmaps and images.

@@

\subsubsection{Sheafs}
\label{sec:math:data:sheaf}
Many types of dynamic visualizations require evaluating sections on different subspaces of \dbase, an a sheaf, denoted $\mathcal{O}$ provides a way to do so. A sheaf is a mathematical structure for defining collections of objects\cite{ghristElementaryAppliedTopology2014, ghristHomologicalAlgebraData2018,urbanikBriefIntroductionSchemes} on mathematical spaces. On the fiber bundle \dtotal, we can describe a sheaf as the collection of local sections $\iota^{*}\dsection$
\begin{equation}
    \label{eq:sheaf}
    \begin{tikzcd}[column sep=huge]
        \iota^*\dtotal \arrow[d, "\pi"'] \arrow[r, "\iota^*", hook]             & \dtotal \arrow[d, "\pi"']                  \\
        U \arrow[r, "\iota", hook] \arrow[u, "\iota^{*}\dsection"', bend right] & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
which are sections of \dtotal\ pulled back over local neighborhood $U \subset \dtotal$ via the inclusion map \(\iota:\dtotal\rightarrow U\). The collation of sections enabled by sheafs is necessary for navigation techniques such as pan and zoom\cite{NekrasovskiEvaluationPanZoom2006} and dynamically updated visualizations such as sliding windows\cite{crouchDynamicGraphsSlidingwindow2013,chuTimeSeriesSegmentation1995}.

@@

\subsection{Graphic Space \gtotal}
\label{sec:math:graphic} 
To establish that the artist is structure preserving map from data \dtotal\ to graphic \gtotal\, we construct a graphic bundle so that we can define \textit{equivariance} in terms of maps on the fiber spaces and \textit{continuity} in terms of maps on the base space. As with the data, we can represent the target graphic as a section \gsection\ of a bundle  $(\gtotal, \gbase, \pi, \gfiber)$. 
\begin{equation}
    \begin{tikzcd}[ampersand replacement=\&]
        \gfiber \arrow[r, hook] \& \gtotal \arrow[d, "\pi"'] \\
                          \& \gbase \arrow[u, "\gsection"', bend right]
    \end{tikzcd}
  \end{equation}
The graphic bundle \gtotal\ consists of a base \gbase (\autoref{sec:math:graphic:fiber}) that is a thickened form of \dbase\, a fiber \gfiber (\autoref{sec:math:graphic:base}) that is an idealized display space, and sections \gsection (\autoref{sec:math:graphic:section}) that encode a graphic where the visual characteristics are fully specified.

\subsubsection{Idealized Display \gfiber}
\label{sec:math:graphic:fiber}
To fully specify the visual characteristics of the image, we construct a fiber \gfiber\ that is an infinite resolution version of the target space. Typically \gtotal\ is trivial and therefore sections can be thought of as mappings into \gfiber. In this work, we assume a 2D opaque image $\gfiber=\reals^5$ with elements 
\begin{equation*}
(x,\, y,\, r,\, g,\, b) \in \gfiber
\end{equation*}
such that a rendered graphic only consists of 2D position and color. To support overplotting and transparency, the fiber could be $\gfiber=\reals^{7}$ such that $(x, y, z, r, g, b, a) \in \gfiber$ specifies the target display. By abstracting the target display space as \gfiber, the model can support different targets, such as a 2D screen or 3D printer.

\subsubsection{Continuity of the Graphic \gbase} 
\label{sec:math:graphic:base}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/retraction_maps.png}
    \caption{For a visualization component to preserve continuity, it must have a continuous surjective map $\vindex: \gbase\rightarrow\dbase$ from graphic continuity to data continuity. The scatter and line graphic base spaces \gbase\ have one more dimension of continuity than \dbase\ so that \gbase\ can encode physical aspects of the glyph, such as shape (a circle) or thickness. The image has the same dimension in \gbase\ as in \dbase.}
    \label{fig:math:graphic:retraction:map}
\end{figure}
To establish that a visualization component preserves continuity, we propose that their must be a continuous map $\vindex: \gbase\rightarrow\dbase$ from the graphic base space to the data space. For example, consider a \gbase\ that is mapped to the region of a 2D display space that represents \dbase. For some visualizations, \dbase\ may be lower dimension than \gbase. For example, a point that is 0D in \dbase\ cannot be represented on screen unless it is thickened to 2D to encode the connectivity of the pixels that visually represent the point. This thickening is often not necessary when the dimensionality of \dbase\ matches the dimensionality of the target space, for example if \dbase\ is 2D and the display is a 2D screen. We introduce \gbase\ to thicken \dbase\ in a way which preserves the structure of \dbase.

Formally, we require that \dbase\ be a deformation retract\cite{RetractionTopology2020} of \gbase\ so that \dbase\ and \gbase\ have the same homotopy, meaning there is a continuous map from \gbase\ to \dbase\cite{weissteinHomotopy}. The surjective map $\vindex: \gbase \rightarrow \dbase$ 
\begin{equation}
    \begin{tikzcd}
        \dtotal \arrow[d, "\pi"'] & \gtotal \arrow[d, "\pi"'] \\
        \dbase                   & \gbase \arrow[l, "\vindex"']
    \end{tikzcd}
\end{equation}
goes from region $\gbasepoint \in \gbase_{\dbasepoint}$ to its associated point $\gbasepoint$. This means that if $\vindex(\gbasepoint) = \dbasepoint$, the record at \dbasepoint\ is copied over the region \gbasepoint\ such that $\dsection(\dbasepoint)=\vindex^*\dsection(\gbasepoint)$ where $\vindex^*\dsection(\gbasepoint)$  is \dsection\ pulled back over \gbase. When \dbase\ is discrete points and the graphic is a scatter plot, each point $\dbasepoint \in \dbase$ corresponds to a 2D disk $\gbase_{\dbasepoint}$ as shown in \autoref{fig:math:graphic:retraction:map}. In the case of 1D continuous data and a line plot, the region \gy\ over a point $\gx_i$ specifies the thickness of the line in \gbase\ for the corresponding \dsection\ on \dbasepoint. The image has the same dimensions in data space and graphic space such that no extra dimensions are needed in \gbase.

@@

\subsubsection{Graphic \gsection}
\label{sec:graphic_section}
The section $\gsection: \gbase \rightarrow \gtotal$ is the graphic in an idealizes prerender space and also acts as a specification for rendering the graphic to an image. It is sufficient to sketch out how an arbitrary pixel would be rendered, where a pixel $p$ in a real display corresponds to a region $\gbase_p$ in the idealized display. To determine the color of the pixel, we aggregate the color values over the region via integration:
\begin{align*}
    \label{eq:math:graphic:section:color}
    r_p &= \iint\limits_{S_p} \rho_r(s)ds^{2}\\
    g_p &= \iint\limits_{S_p} \rho_g(s)ds^{2}\\
    b_p &= \iint\limits_{S_p} \rho_b(s)ds^{2}
\end{align*}
For a 2D screen, the pixel is defined as a region $p=\left[y_{top}, y_{bottom}, x_{right}, x_{left}\right]$ of the rendered graphic. Since the x and y in $p$ are in the same coordinate system as the x and y components of \gfiber\,  the inverse map of the bounding box $\gbase_{p} ={\gsection_{xy}}^{-1}(p)$ is a region $\gbase_p \subset \gbase$. The color is the result of the integration over  $\gbase_p$.

\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/render.png}
    \caption{To render a graphic, a pixel $p$ is selected in the display space, which is defined in the same coordinates as the x and y components in \gfiber\ via the renderer. In \gtotal\, the inverse mapping ${\gsection_{xy}}^(p)$ returns a region $\gbase_{p} \subset \gbase$. $\gsection(\gbase_{p})$ returns a set of points $(x,\,y,\,r,\,g,\,b) \in \gfiber$ that lie over $\gbase_{p}$. The integral over the $(r,\,g,\,b)$ pixels specifies that the pixel should be green}
    \label{fig:math:graphic:rho:lookup}
\end{figure}

As shown in \autoref{fig:math:graphic:rho:lookup}, a pixel $p$ in the output space, drawn in yellow, is selected and mapped, via the renderer, into a region on \gtotal.  The region on \gtotal\ corresponds to a region $\gbase_p \subset \gbase$ via the inverse mapping  ${\gsection_{xy}}^(p)$. The base space \gbase\ is an annulus to match the topology of the graphic idealized in \gtotal. The section  $\gsection(\gbase_{p})$  then maps into the fiber \gfiber\ over $\gbase_p$ to obtain the set of points in \gfiber, here represented as a table, that correspond to that section. The integral over the pixel components of this set of points in the fiber yields \{3, 135, 60\} the actual color of the pixel. In general, \gsection\ is an abstraction of rendering. In very broad strokes \gsection\  can be a specification such as PDF\cite{bienz1993portable}, SVG\cite{quintScalable2003}, or an openGL scene graph\cite{CarsonOpenGL1997}. Alternatively, \gsection\ can be a rendering engine such as cairo\cite{CairographicsOrg} or AGG\cite{shemanarevAntiGrainGeometry}. Implementation of \gsection\ is out of scope for this work,

\subsection{Artist}
\label{sec:artist}
We propose that the transformation from data to visual representation can be described as a structure preserving map from one topological space to another. We name this map the artist as that is the analogous part of the  Matplotlib\cite{hunterArchitectureOpenSource} architecture that builds visual elements. The topological artist \vartist\ is a monoid equivariant sheaf map from the sheaf on a data bundle \dtotal\ which is $\mathcal{O}(\dtotal)$ to the sheaf on the graphic bundle \gtotal, $\mathcal{O}(\gtotal)$. 
\begin{equation}
    \vartist: \mathcal{O}(\dtotal) \rightarrow \mathcal{O}(\gtotal)
\end{equation}
Sheafs are a mathematical object with restriction maps that define how to glue \dsection over local neighborhoods $U \subseteq \dbase$,  discussed in \autoref{sec:artist_sec:data_sheaf_stalk}, such that the \vartist\ maps are consistent over continuous regions of \dbase. While \vartist can usually construct graphical elements solely with the data in \dsection, some visualizations, such as line, may also need some finite number $n$ of derivatives, which is captured by the jet bundle $\mathcal{J}^n$ \cite{JetBundle2020,musilovaCalculusVariationsJet2016} with $\mathcal{J}^{0}(\dtotal)=\dtotal$. In this work, we at most need $\mathcal{J}^{2}(\dtotal)$ which is the value at \dsection\ and its first and second derivatives; therefore the artist takes as input the jet bundle $\dtotal^{\prime}=\mathcal{J}^{2}(\dtotal)$.

@@

\begin{equation}
    \label{eq:artist_diagram}
    \begin{tikzcd}
        \dtotal^{\prime} \arrow[r, "\vchannel"] \arrow[rd, "\pi"'] & \vtotal \arrow[d, "\pi"] & \vindex^*\vtotal \arrow[r, "\vmark"] \arrow[d, "\vindex^*\pi"'] \arrow[l, "\vindex^*"'] & \gtotal \arrow[ld, "\pi"] \\
                                              & \dbase                  & \gbase \arrow[l, "\vindex"']                                              &                    
        \end{tikzcd}
\end{equation}
where the input can be point wise $\dsection(\dbasepoint)\mid \dbasepoint \in \dbase$. The encoders $\vchannel:\dtotal^{\prime} \rightarrow \vtotal$ convert the data components to visual components(\autoref{sec:graphic_base}). The continuity map $\vindex:\gbase \rightarrow \dbase$ then pulls back the visual bundle \vtotal\ over \gbase (\autoref{sec:artist_nu}). Then the assembly function $\vmark: \vtotalpull \rightarrow \gtotal$ composites the fiber components of \vtotalpull\ into a graphic in \gtotal (\autoref{sec:artist_q}). This functional decomposition of the visualization artist facilitates building reusable components at each stage of the transformation because the equivariance constraints are defined on \vchannel, \vmark, and \vindex.

\subsubsection {Visual Fiber Bundle \vtotal}
We introduce a visual bundle \vtotal\ to store the visual representations the artist needs to assemble into a graphic. The visual bundle (\vtotal, \dbase, $\pi$, \vfiber) has section $\vsection: \vtotal \rightarrow \dbase$ that resolves to a visual variable in the fiber \vfiber. The visual bundle \vtotal\ is the latent space of possible parameters of a visualization type, such as a scatter or line plot. We define \vfiber\ in terms of the parameters of a visualization libraries compositing functions; for example \autoref{tab:mpl_visual_variable_fiber} is a sample of the fiber space for Matplotlib \cite{hunterMatplotlib2DGraphics2007}.

@@

\subsubsection{Visual Encoders \vchannel}
\label{sec:artist_nu}
%% equivariant maps 
\begin{figure}[H]
    \includegraphics[width=\textwidth]{figures/math/partial_fixed.png}
    \caption{In this artist, \vchannel\ maps the strings to the emojis. This \vchannel\ is equivariant because the monoid actions (which are represented by the colored arrows) are the same on both the \dsection\ input and \vsection\ output sets.}
    \label{fig:artist_nu}
\end{figure}
As introduced in \autoref{sec:intro_viz}, there are many ways to visually represent data components. We define the visual transformers \vchannel\ 
\begin{equation}
    \label{eq:nu_expanded}
    \{\vchannel_{0}, \ldots, \vchannel_{n}\}: \{\dsection_{0}, \ldots, \dsection_{n}\} \mapsto \{\vsection_{0}, \ldots, \vsection_{n}\}
\end{equation}
as the set of equivariant maps $\vchannel_i: \dsection_i \mapsto \vsection_i$. Given $\monoid_i$ is the monoid action on $\dtotal_i$ and that there is a monoid ${\monoid_i}^{\prime}$ on $\vtotal_i$, then there is a monoid homomorphism from $\varphi:\monoid_i \rightarrow {\monoid_i}^{\prime}$ that \vchannel\ must preserve. As mentioned in \autoref{sec:data_monoid}, we choose monoid actions as the basis for equivariance because they define the structure on the fiber components.

@@

On example of an equivariant \vchannel\ is illustrated in \autoref{fig:artist_nu}, which is a mapping from \textbf{Strings} to symbols. The data is an example of a Steven's nominal measurement set, which is defined as having on it permutation group actions 
\begin{equation}
    \text{ if } \delement_1 \neq \delement_2 \text{ then } \vchannel (\delement_1) \neq\vchannel(\delement_2)
\end{equation}
such that shuffling the words must have an equivalent shuffle of the symbols they are mapped to. This is illustrated in the identical actions, represented by the colored arrows, on the words and emojis. To preserve ordinal and partial order monoid actions, \vchannel\ must be a monotonic function such that given $\delement_1, \delement_2 \in \dtotal_{i}$,
\begin{equation}
\text{ if } \delement_1 \leq \delement_2 \text{ then } \vchannel (\delement_1) \leq \vchannel(\delement_2)
\end{equation}
the visual encodings must also have some sort of ordering. For interval scale data, \vchannel\ is equivariant under translation monoid actions if 
\begin{equation}
\vchannel (x + c) = \vchannel(x) + c
\end{equation}
while for ratio data, there must be equivalent scaling\cite{weissteinSimilarityTransformation} 
\begin{equation}
\vchannel(xc) = \vchannel(x)*c 
\end{equation}
We therefore can test if a \vchannel\ is equivariant by testing the actions under which is must  commute. For example, we define a transform $\vchannel_{i}(x) = .5$ on interval data. This means it must commute under translation, for example $t(x) = x+2$. Testing this constraint 
\begin{align}
    \vchannel(t(\delement + 2)) & \overset{?}{=} \vchannel(\delement) + 2\\
    .5 &\neq .5 + 2
\end{align}
we find that the \vchannel\ defined here does not commute and is therefore invalid. The constraints on \vchannel\ can be embedded into our artist such that the \vchannel\ functions can test for equivariance and also provide guidance on constructing new \vchannel\ functions.

@@

As shown in \autoref{fig:artist_q}, the assembly function \vmark\ combines the fiber $\dfiber_i$ wise \vchannel\ transforms into a graphic in \gtotal. Together, \vchannel\ and \vmark\ are a map-reduce operation: map the data into their visual encodings, reduce the encodings into a graphic. As with \vchannel\, the constraint on \vmark\ is that for every monoid action on the input \vsection\, there is corresponding monoid action on the output \gsection.

@@

The glyph in \autoref{fig:artist_mark_change} has the following characteristics \vfiber\ specified by  $(xpos,\, ypos,\, color,\, thickness)$ such that one section is $\vsection=(0,0,0,1)$ and $\vmark(\vsection) = \gsection$ generates a piece of the thin hollow circle. The equivariance constraint on \vmark\ is that the action $m=(e, e, e, x+2)$, where e is identity, translates \vsection\ to  $\vsection^{\prime}=(e,e,e,3)$. The corresponding action on \gsection\ causes $\vmark(\vsection^{\prime})$ to be the thicker circle in \autoref{fig:artist_mark_change}.

\subsubsection{Assembly \vmark}
In this section we formulate the minimal Q that will generate distinguishable graphical marks: non-overlapping scatter points, a non-infinitely thin line, and an image. 
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/scatter_with_s.png}
    \caption{The data is discrete points (temperature, time). Via \vchannel\ these are converted to (xpos, ypos) and pulled over discrete \gbase. These values are then used to parameterize \gsection\ which returns a color based on the parameters (xpos,ypos) and position $\alpha, \beta$ on $\gbase_{\dbasepoint}$ that \gsection\ is evaluated on. 
    }
    \label{fig:artist_scatter}
\end{figure}
The scatter plot in \autoref{fig:artist_scatter} can be defined as $\vmark(xpos, ypos)(\alpha, \beta)$ where color $\gsection_{RGB} = (0,0,0)$ is defined as part of \vmark\ and $\gbasepoint=(\alpha, \beta)$ defines the region on \gbase. The position of this swatch of color can be computed relative to the location on the disc $\gbase_{\dbasepoint}$ as shown in \autoref{fig:artist_scatter}:
\begin{align}
x &= size *\alpha \cos(\beta) + xpos \\
y &= size *\alpha \sin(\beta) + ypos
\end{align}

such that $\gsection(\gbasepoint) = (x, y, 0, 0, 0)$ colors the point (x,y) black. Here $size$ can either be defined inside \vmark\ or it could also be a parameter in \vtotal\ that is passed along with (xpos, vpos). As seen in \autoref{fig:artist_scatter}, a scatter has a direct mapping from a region on $\gbase_{k}$ to its corresponding \dbasepoint.
